--- 
title: "Arbuscular mycorrhizal fungi colonization package"
author: "Marco Chiapello"
output:  
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Ramf colonization analysis} 
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r env, echo=FALSE, warnings=FALSE}
library("Ramf")
library("BiocStyle")
```
### Introduction

The package is meant to help scientists to compute the degree of arbuscular
mycorrhizal colonization, producing tables, robust statistics and different
kind of plots.

All the functions of the package work independently of the colonization scoring system.

#### Grid scoring system

The data should be provided in a **csv table** like the follow one:

|  samples  | replicates | Total | Hyphopodia | IntrHyphae | Arbuscule | Vesicle |
|:---------:|:----------:|:-----:|:----------:|:----------:|:---------:|:-------:|
|     WT    |      A     |   30  |      3     |     30     |     30    |    10   |
|     WT    |      A     |   24  |      5     |     24     |     24    |    7    |
|     WT    |      B     |   20  |      2     |     20     |     20    |    5    |
|     WT    |      B     |   15  |      2     |     15     |     15    |    3    |
| cyclops-3 |      A     |   18  |      6     |     18     |     9     |    0    |
| cyclops-3 |      A     |   16  |     10     |     10     |     0     |    0    |
| cyclops-3 |      B     |   20  |     12     |     20     |     0     |    0    |
| cyclops-3 |      B     |   14  |      4     |     14     |     0     |    0    |
|    ...    |     ...    |  ...  |     ...    |     ...    |    ...    |   ...   |

The table contains:

- 7 columns with **specific headers**

- The first 2 columns (samples and replicates) must be present

- The last 5 columns (Total, Hyphopodia, IntHyphae, Arbuscule and Vesicle) could not be all present

- If the headers are not the same of the ones in the example table the software will not be able to read the table

- As many row as you need

- The **first element of column "samples" will be used as control** for the statistical analysis (in this example "WT" will be used as control)

#### Trouvelot scoring system

| samples | replicates | scoring |
|:-------:|:----------:|:-------:|
|  Lys11  |      A     |   1A3   |
|  Lys11  |      A     |   5A3   |
|  Lys11  |      A     |   0A0   |
|  Lys11  |      B     |   3A2   |
|  Lys11  |      B     |   3A3   |
|  Lys11  |      B     |   4A3   |
|   GiFu  |      A     |   0A0   |
|   GiFu  |      A     |   0A0   |
|   GiFu  |      A     |   1A3   |
|   GiFu  |      B     |   3A3   |
|   GiFu  |      B     |   0A0   |
|   GiFu  |      B     |   0A0   |
|   ...   |     ...    |   ...   |

### Loading the data 

To load data in, use the follow command: 

>readData(file.csv, type = c("grid", "trouvelot"))

#### Example
```{r loadData}
# Grid
f <- dir(system.file("extdata", package = "Ramf"), full.names = TRUE, pattern = "grid.csv")
gr <- readData(f, type = "grid")
# Grid incomplete
f <- dir(system.file("extdata", package = "Ramf"), full.names = TRUE, pattern = "grid_incomplete.csv")
gri <- readData(f, type = "grid")
# Trouvelot
f <- dir(system.file("extdata", package = "Ramf"), full.names = TRUE, pattern = "trouvelot.csv")
tr <- readData(f, type = "trouvelot") 
```

It is very important the order of the samples in the csv file. The first sample will
be considered the control.


### Summarize data

To summarise the data and export the scoring results in a tabular format use the
follow command:

> am_summary


#### Example
```{r summary}
grs <- am_summary(gr)
gris <- am_summary(gri)
trs <- am_summary(tr)
grs
gris
trs
```

### Statistical analysis

The statistical analysis is performed using the `kruskal` function from `agricolae` package.
To read more about the statatical framework, refer to this [link](https://en.wikipedia.org/wiki/Kruskal%E2%80%93Wallis_one-way_analysis_of_variance).

> am_stat


#### Example
```{r stat}
grst <- am_stat(gr)
grist <- am_stat(gri)
trst <- am_stat(tr)
grst
grist
trst
```

By default, the correction method is "none", but it is possible to select between:
Bonferroni, Holm, Hochberg, Bonferroni, Benjamini-Hochberg, Benjamini-Yekutieli or fdr 
adjustment.
 For more details refer to [agricolae package](https://cran.r-project.org/web/packages/agricolae/index.html).

#### Example
```{r stat2}
am_stat(gr, method = "BH")
am_stat(gri, method = "BH")
am_stat(tr, method = "fdr")
```

### Plot data

To summarise the data in a graphical way the package provide 3 plot types:

- barplot

  >am_barplot

- boxplot

  >am_boxplot

- dotplot

  >am_dotplot

It is possible to plot a simple plot.

#### Example
```{r plot}
am_barplot(gr)
am_boxplot(gr)
am_dotplot(gr)
am_barplot(tr)
am_boxplot(tr)
am_dotplot(tr)
```

The plot fuctions take a set of option to customize your plot.

#### Example

The plot can use different color palettes. We advise to use [Colorbrewer](http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3) or
[Colorgorical](http://vrl.cs.brown.edu/color).
```{r plot2}
am_barplot(gr, cbPalette = c('#f7f7f7', '#d9d9d9', '#bdbdbd', '#969696'))
am_dotplot(gr, cbPalette = rep("black", 4))
am_barplot(gr, cbPalette = c('#ca0020', '#f4a582', '#92c5de', '#0571b0'))
am_boxplot(gr, cbPalette = c("#b4ddd4", "#02531d", "#34debb", "#d10f55"))
am_barplot(gri, cbPalette = c('#f7f7f7', '#d9d9d9', '#bdbdbd', '#969696'))
am_dotplot(gri, cbPalette = rep("black", 4))
am_barplot(gri, cbPalette = c('#ca0020', '#f4a582', '#92c5de', '#0571b0'))
am_boxplot(gri, cbPalette = c("#b4ddd4", "#02531d", "#34debb", "#d10f55"))
```

The title of the plot can be changed
```{r title}
am_barplot(gr, main = "Grid plot")
am_barplot(gri, main = "Grid incomplete plot")
am_barplot(tr, main = "Trouvelot plot")
```

The plotting system is based on [ggplot2](http://ggplot2.org/), 
so is possible to change some parameters

```{r title.position}
library(ggplot2)
am_barplot(gr, main = "Grid plot") + theme(plot.title = element_text(hjust = .5))
am_barplot(tr, main = "Trouvelot plot") + theme(plot.title = element_text(hjust = 1))
am_barplot(gr, main = "Grid plot") + theme(panel.grid.major.y = element_line(size = 3, colour = "grey80"))
am_barplot(gr, main = "Grid plot") + theme(text=element_text(family="Avenir"))
am_barplot(gr, main = "Grid plot") + theme(axis.text = element_text(size = 18))
```

It is also possible to include the statistical results under the plot with the 
appropriate alpha

```{r stat.plot}
am_barplot(gr, annot = "asterisks", alpha = 0.05)
am_barplot(gri, annot = "asterisks", alpha = 0.05)
am_dotplot(tr, annot = "asterisks", alpha = 0.01)
```

The asterisks below the plot show which sample is significant different from the control. 
The control is always the first element of the list.

It is also possible to include the statistical correction for the pvalues.

```{r stat.plot2}
am_barplot(gr, annot = "asterisks", method = "BH")
```

The asterisks allow to grafically dispay the statistical difference between the
control sample and the other samples. If you want show the statistical difference
between all the samples, you can add "letters" to the plot. Letters group the
samples accordingly to they statistical difference.

```{r stat.plot3}
am_barplot(gr, annot = "letters", alpha = 0.05)
am_barplot(gri, annot = "letters", alpha = 0.05)
am_dotplot(tr, annot = "letters", alpha = 0.01)
```


### Save data

To export the data, a "save data" function is provided. 

- **Summary data**: the function exports both the summary per replicate and the summary per sample in two distinct csv files.

- **Statistical data**: the function exports a csv file

- **Plots**: the function exports the plots and useful options can be used to export a pubblication ready image.


#### Example
```{r save}
# Save the summary
am_save(grs) # Two files will be saved: Summary_per_Replicate.csv and Summary_per_Sample.csv
am_save(trs, "Trouvelot") # Two files will be saved: Trouvelot_per_Replicate.csv and Trouvelot_per_Sample.csv

# Save the statisticas
am_save(am_stat(gr)) # One file called Summary_Stat.csv will be saved
am_save(am_stat(tr, method = "BH"), "Trouvelot") # One file called Trouvelot_stat.csv will be saved

# Save the plots
am_save(am_barplot(gr), "barplor_grid.pdf")
p1 <- am_barplot(gr, cbPalette = c('#f7f7f7', '#d9d9d9', '#bdbdbd', '#969696'), main = "Trouvelot", annot = "letters", method = "BH")
am_save(p1, "dotplor_trouvelot.pdf", width = 21, height = 21, units = "cm", dpi = 300)

# Save jpg plot
am_save(p1, "dotplor_trouvelot.jpg")
```

### Time series analysis

The package contains functions able to deal with time course data.

#### Grid scoring system for time point data

The data should be provided in a **csv table** like the follow one:

|  samples  | replicates | time | Total | Hyphopodia | IntrHyphae | Arbuscule | Vesicle |
|:---------:|:----------:|:----:|:-----:|:----------:|:----------:|:---------:|:-------:|
|     WT    |      A     |   4  |   30  |      3     |     30     |     30    |    10   |
|     WT    |      A     |   4  |   24  |      5     |     24     |     24    |    7    |
|     WT    |      B     |   4  |   20  |      2     |     20     |     20    |    5    |
|     WT    |      B     |   5  |   15  |      2     |     15     |     15    |    3    |
| cyclops-3 |      A     |   5  |   18  |      6     |     18     |     9     |    0    |
| cyclops-3 |      A     |   5  |   16  |     10     |     10     |     0     |    0    |
| cyclops-3 |      B     |   6  |   20  |     12     |     20     |     0     |    0    |
| cyclops-3 |      B     |   6  |   14  |      4     |     14     |     0     |    0    |
|    ...    |     ...    | ...  |  ...  |     ...    |     ...    |    ...    |   ...   |

The table contains:

- 8 columns with **specific headers**

- The first 3 columns (samples, replicates and time) must be present

- The last 5 columns (Total, Hyphopodia, IntHyphae, Arbuscule and Vesicle) could not be all present

- If the headers are not the same of the ones in the example table the software will not be able to read the table

- As many row as you need

- The **first element of column "samples" will be used as control** for the statistical analysis (in this example "WT" will be used as control)

### Loading time point  data 

To load data in, use the follow command: 

>readDataTime(file.csv, type = c("grid", "trouvelot"))

#### Example
```{r loadDataTime}
# Grid
f <- dir(system.file("extdata", package = "Ramf"), full.names = TRUE, pattern = "gridTime.csv")
tgr <- readDataTime(f, type = "grid")
```

It is very important the order of the samples in the csv file. The first sample will
be considered the control.


### Summarize data

To summarise the data and export the scoring results in a tabular format use the
follow command:

> am_summary


#### Example
```{r summaryTime}
tgrs <- am_summary(tgr)
tgrs
```

### Statistical analysis

The statistical analysis is performed using the `kruskal` function from `agricolae` package.
To read more about the statatical framework, refer to this [link](https://en.wikipedia.org/wiki/Kruskal%E2%80%93Wallis_one-way_analysis_of_variance).

> am_stat


#### Example
```{r statTime}
tgrst <- am_stat(tgr)
tgrst
```

By default, the correction method is "none", but it is possible to select between:
Bonferroni, Holm, Hochberg, Bonferroni, Benjamini-Hochberg, Benjamini-Yekutieli or fdr 
adjustment.
 For more details refer to [agricolae package](https://cran.r-project.org/web/packages/agricolae/index.html).

#### Example
```{r stat2Time}
am_stat(tgr, method = "BH")
```

The function reports the statistical analysis between the samples at the same time point and the statistical
analysis of the same sample at different time points.

### Plot data

To summarise the data in a graphical way the package provide 3 plot types:

- barplot

  >am_barplot

- boxplot

  >am_boxplot

- dotplot

  >am_dotplot

It is possible to plot a simple plot.

#### Example
```{r plotTime}
am_barplot(tgr)
am_boxplot(tgr)
am_dotplot(tgr)
```

The plot fuctions for time points take the same set of option to customize your plot as the previous plots.

It is also possible to include the statistical results under the plot with the 
appropriate alpha. In time point data analysis the "asterisks" option does not work. The "letter" option
works across the same feature for all the time points. So, you can compare the "Total" for different samples,
 but you can not compare different features at different time points (eg, the comparison between "Total" HP0 after 4 days
 and "Arbuscule" HP0 after 4 days is meaningless).

```{r stat.plotTime}
am_barplot(tgr, annot = "letter", alpha = 0.05)
am_barplot(tgr, annot = "letter", alpha = 0.01)
am_dotplot(tgr, annot = "letter", method = "fdr")
```

### Save data

To export the data, a "save data" function is provided. 

- **Summary data**: the function exports both the summary per replicate and the summary per sample in two distinct csv files.

- **Statistical data**: the function exports a csv file

- **Plots**: the function exports the plots and useful options can be used to export a pubblication ready image.


#### Example
```{r saveTime}
# Save the summary
am_save(tgrs, "TimePoint") # Two files will be saved

# Save the statisticas
am_save(am_stat(tgr, method = "BH"), "TimePoint") # One file called Summary_Stat.csv will be saved

# Save the plots
am_save(am_barplot(tgr), "barplor_Timegrid.pdf")

# Save jpg plot
tgrb <- am_barplot(tgr)
am_save(tgrb, "RPlot.pdf", unit = "cm", width = 20, height = 20, dpi = 300)
```
