--- 
title: "Arbuscular mycorrhizal fungi colonization" 
author: "Marco Chiapello"
output:  
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Ramf colonization analysis} 
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r env, echo=FALSE, warnings=FALSE}
library("Ramf")
library("BiocStyle")
```
### Introduction

The package is meant to help scientists to compute the degree of arbuscular
mycorrhizal fungi.

All the functions of the package work independently of the colonization scoring system.

#### Grid scoring system

...

#### Trouvelot scoring system

...

### Loading the data 

To load data in, use the follow command: 

>readData

#### Example
```{r loadData}
# Grid
f <- dir(system.file("extdata", package = "Ramf"), full.names = TRUE, pattern = "grid.csv")
gr <- readData(f)
# Trouvelot
f <- dir(system.file("extdata", package = "Ramf"), full.names = TRUE, pattern = "trouvelot.csv")
tr <- readData(f) 

```

### Summarize data

To summarise the data and export the scoring results in a tabular format use the
follow command:

> am_summary


#### Example
```{r summary}
grs <- am_summary(gr)
trs <- am_summary(tr)
grs
trs
```

### Plot data

To summarise the data in a graphical way the package provide 2 plot types:

- barplot

  >am_barplot

- boxplot

  >am_boxplot

#### Example
```{r plot}
p1 <- am_barplot(gr)
p2 <- am_boxplot(gr)
p3 <- am_barplot(tr)
p4 <- am_boxplot(tr)
```

### Save data

To export the summary data, a "save data" function is provided. The function 
exports both the summary per replicate and the summary per sample in two distinct
csv files.

#### Example
```{r save}
am_save(grs)
am_save(trs)
am_save(p1, "barplor_grid.pdf")
am_save(p4, "boxplor_trouvelot.pdf")
```


