--- 
title: "Arbuscular mycorrhizal fungi colonization" 
author: "Marco Chiapello"
output:  
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{Ramf colonization analysis} 
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r env, echo=FALSE, warnings=FALSE}
library("Ramf")
library("BiocStyle")
```
### Introduction

The package is meant to help scientists to compute the degree of arbuscular
mycorrhizal fungi.

All the functions of the package work independently of the colonization scoring system.

#### Grid scoring system

...

#### Trouvelot scoring system

...

### Loading the data 

To load data in, use the follow command: 

>readData

#### Example
```{r loadData}
# Grid
gr <- readData("../inst/extdata/grid.csv") 
# Trouvelot
tr <- readData("../inst/extdata/trouvelot.csv") 

```

### Summarize data

To summarise the data and export the scoring results in a tabular format use the
follow command:

> am_summary


#### Example
```{r summary}
am_summary(gr)
am_summary(tr)
```

### Plot data

To summarise the data in a graphical way the package provide 2 plot types:

- barplot

  >am_barplot

- boxplot

  >am_boxplot

#### Example
```{r plot}
am_barplot(gr)
am_boxplot(gr)
am_barplot(tr)
am_boxplot(tr)
```

